
---
title: "Machine Learning Project"
author: "Jared Allen"
date: "29/08/2019"
output: html_document
---

##Introduction - Weight Lifting Dataset

Data obtained from http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. The goal of this analysis is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict common mistakes made during weightlifting exercises.

Six young health participants were each asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. A relatively light dumbbell was used (1.25kg) to ensure that all participants could easily simulate the mistakes in a safe and controlled manner.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(knitr)
library(kableExtra)

library(PerformanceAnalytics)
library(corrplot)
library(car)
library(doParallel)
library(ggfortify)
library(cluster)
library(plotly)
library(caretEnsemble)
library(dlookr)
```

```{r dataimport, echo=FALSE, message=FALSE, warning=FALSE}
training_raw <- read_csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",col_names=TRUE)
testing <- read_csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",col_names=TRUE)
```

###Prediction of classe variable

###Step 1 - EDA & removal of uninformative variables

An initial look at the data shows a large number of variables which contribute very litle information.

Variables with NA values were removed, along with uninformative variables such as user_name, and timestamp data, to give a reduced featureset upon which to train a model.

```{r EDA, echo=FALSE, message=FALSE, warning=FALSE}
#Remove uninformative
training <- training_raw[sapply(training_raw, function(x) !any(is.na(x)))] 
training <- training[,-1:-7]
heatmap(data.matrix(training[,-53]))
```

The data was split by sensor group (belt, arm, dumbbell, and forearm), and individual random forest models were trained on each set of measurements, as well as a model trained on the entire dataset, with 5-fold cross validation used to assess in sample error for the training set for each model. These are summarised in Appendix A, with 5-fold cross validation showing accuracies ranging from 71.5% to 99.4%.

```{r subsetting, echo=FALSE, message=FALSE}

## apply 5-fold cross validation to the model training

set.seed(875)
train_control <- trainControl(method="cv", number=5)

##sensors <- c("belt", "arm", "dumbbell", "forearm")
dta_train_belt <-
  training %>%
  select(matches("classe|._belt."))

dta_train_arm <-
  training %>%
  select(matches("classe|._arm."))

dta_train_dumbbell <-
  training %>%
  select(matches("classe|._dumbbell."))

dta_train_forearm <-
  training %>%
  select(matches("classe|._forearm."))

```

```{r models, echo=FALSE, message=FALSE}

set.seed(123)
mdl_belt_rf <- train(classe~.,
                   dta_train_belt,
                   trControl=train_control,
                   method = "rf")

set.seed(123)
mdl_arm_rf <- train(classe~.,
                   dta_train_belt,
                   trControl=train_control,
                   method = "rf")

set.seed(123)
mdl_dumbbell_rf <- train(classe~.,
                   dta_train_dumbbell,
                   trControl=train_control,
                   method = "rf")

set.seed(123)
mdl_forearm_rf <- train(classe~.,
                   dta_train_forearm,
                   trControl=train_control,
                   method = "rf")

set.seed(123)
mdl_full_rf <- train(classe~.,
                   training,
                   trControl=train_control,
                   method = "rf")
```

```{r modelensemble, echo=FALSE, message=FALSE, warning=FALSE}

tconfirm <- as.factor(training$classe)
tp_belt_rf <- predict(mdl_belt_rf, newdata=training)
tp_arm_rf <- predict(mdl_arm_rf, newdata=training)
tp_dumbbell_rf <- predict(mdl_dumbbell_rf, newdata=training)
tp_forearm_rf <- predict(mdl_forearm_rf, newdata=training)
tp_full_rf <- predict(mdl_full_rf, newdata=training)


train_ensemble <- cbind.data.frame(classe=tconfirm,belt_rf=tp_belt_rf,arm_rf=tp_arm_rf,dumbbell_rf=tp_dumbbell_rf,forearm_rf=tp_forearm_rf,full_rf=tp_full_rf)

set.seed(897)

mdl_ensemble <- train(classe~.,
                   train_ensemble,
                   trControl=train_control,
                   method = "nb")

```

```{r OOSAccuracy}

accuracysummary <- data.frame(
  Models = c("Arm_RF","Belt_RF","Dumbbell_RF","Forearm_RF","Full_RF","Ensemble_NB"),
  Mean_OOS_Accuracy = c(mean(mdl_arm_rf$resample$Accuracy),
                        mean(mdl_belt_rf$resample$Accuracy),
                        mean(mdl_dumbbell_rf$resample$Accuracy),
                        mean(mdl_forearm_rf$resample$Accuracy),
                        mean(mdl_full_rf$resample$Accuracy),
                        mean(mdl_ensemble$resample$Accuracy))
)

kable(accuracysummary, caption="Summary of mean out of sample accuracy during 5-fold cross-validation") %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

```

A final ensemble model was assembled using the predictions of the 5 random forest models to train a naive bayes classifier to predict final class.

This model gave perfect accuracy within 5-fold cross validation.

## APPENDIX A: Confusion Matrices for Trained Models

```{r beltmodel}
print("Random forest trained on Belt sensor data")
confusionMatrix(mdl_belt_rf, norm = "none", dnn=c("Prediction","Reference"))

```

```{r armmodel}
print("Random forest trained on Arm sensor data")
confusionMatrix(mdl_arm_rf, norm = "none", dnn=c("Prediction","Reference"))

```

```{r dumbbellmodel}
print("Random forest trained on Dumbbell sensor data")
confusionMatrix(mdl_dumbbell_rf, norm = "none", dnn=c("Prediction","Reference"))

```

```{r forearmmodel}
print("Random forest trained on Forearm sensor data")
confusionMatrix(mdl_forearm_rf, norm = "none", dnn=c("Prediction","Reference"))

```

```{r fullmodel}
print("Random forest trained on combination of all sensor data")
confusionMatrix(mdl_full_rf, norm = "none", dnn=c("Prediction","Reference"))

```



```{r confmat, echo=FALSE, message=FALSE, warning=FALSE}

print("Ensemble Naive Bayes MOdel trained on predictions from previous models")

confusionMatrix(mdl_ensemble, norm = "none", dnn=c("Prediction","Reference"))

```

```{r confmat2, echo=FALSE, message=FALSE, warning=FALSE}

test_belt_rf <- predict(mdl_belt_rf, newdata=testing)
test_arm_rf <- predict(mdl_arm_rf, newdata=testing)
test_dumbbell_rf <- predict(mdl_dumbbell_rf, newdata=testing)
test_forearm_rf <- predict(mdl_forearm_rf, newdata=testing)
test_full_rf <- predict(mdl_full_rf, newdata=testing)
test_ensemble <- cbind.data.frame(belt_rf=test_belt_rf,arm_rf=test_arm_rf,dumbbell_rf=test_dumbbell_rf,forearm_rf=test_forearm_rf,full_rf=test_full_rf)

trainingpred <- predict(mdl_ensemble, test_ensemble)
trainingpred

```