
---
title: "Machine Learning Project"
author: "Jared Allen"
date: "29/08/2019"
output: html_document
---

##Introduction

Data obtained from http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

What you should submit
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

Peer Review Portion
Your submission for the Peer Review portion should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).

## Weight Lifting Exercises Dataset

On-body sensing schema

This human activity recognition research has traditionally focused on discriminating between different activities, i.e. to predict "which" activity was performed at a specific point in time (like with the Daily Living Activities dataset above). The approach we propose for the Weight Lifting Exercises dataset is to investigate "how (well)" an activity was performed by the wearer. The "how (well)" investigation has only received little attention so far, even though it potentially provides useful information for a large variety of applications,such as sports training.

In this work (see the paper) we first define quality of execution and investigate three aspects that pertain to qualitative activity recognition: the problem of specifying correct execution, the automatic and robust detection of execution mistakes, and how to provide feedback on the quality of execution to the user. We tried out an on-body sensing approach (dataset here), but also an "ambient sensing approach" (by using Microsoft Kinect - dataset still unavailable)

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. We made sure that all participants could easily simulate the mistakes in a safe and controlled manner by using a relatively light dumbbell (1.25kg).


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)

library(PerformanceAnalytics)
library(corrplot)
library(car)
library(doParallel)
library(ggfortify)
library(cluster)
library(plotly)
library(caretEnsemble)
```

```{r dataimport, echo=FALSE, message=FALSE, warning=FALSE}
training_raw <- read_csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",col_names=TRUE)
testing <- read_csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",col_names=TRUE)
```

Prediction of classe variable

Step 1 - EDA & removal of uninformative variables

```{r EDA, echo=FALSE, message=FALSE, warning=FALSE}
#Remove uninformative
training <- training_raw[sapply(training_raw, function(x) !any(is.na(x)))] 
training <- training[,-1:-7]

```


```{r subsetting, echo=FALSE, message=FALSE}

##split 'training' data into training and test sets, leave 'testing' as hold out validation set

set.seed(875)
train_control <- trainControl(method="cv", number=5)

##sensors <- c("belt", "arm", "dumbbell", "forearm")
dta_train_belt <-
  training %>%
  select(matches("classe|._belt."))

dta_train_arm <-
  training %>%
  select(matches("classe|._arm."))

dta_train_dumbbell <-
  training %>%
  select(matches("classe|._dumbbell."))

dta_train_forearm <-
  training %>%
  select(matches("classe|._forearm."))

```

```{r models2, echo=FALSE, message=FALSE}

set.seed(123)
mdl_belt_rf <- train(classe~.,
                   dta_train_belt,
                   trControl=train_control,
                   method = "rf")

mdl_arm_rf <- train(classe~.,
                   dta_train_belt,
                   trControl=train_control,
                   method = "rf")

set.seed(123)
mdl_dumbbell_rf <- train(classe~.,
                   dta_train_dumbbell,
                   trControl=train_control,
                   method = "rf")

set.seed(123)
mdl_forearm_rf <- train(classe~.,
                   dta_train_forearm,
                   trControl=train_control,
                   method = "rf")
```

```{r beltmodel}
print(mdl_belt_rf)
confusionMatrix(mdl_belt_rf, norm = "none", dnn=c("Prediction","Reference"))

```

```{r beltmodel}
print(mdl_arm_rf)
confusionMatrix(mdl_arm_rf, norm = "none", dnn=c("Prediction","Reference"))

```

```{r beltmodel}
print(mdl_dumbbell_rf)
confusionMatrix(mdl_dumbbell_rf, norm = "none", dnn=c("Prediction","Reference"))

```

```{r beltmodel}
print(mdl_forearm_rf)
confusionMatrix(mdl_forearm_rf, norm = "none", dnn=c("Prediction","Reference"))

```

```{r modelensemble, echo=FALSE, message=FALSE, warning=FALSE}

tconfirm <- as.factor(training$classe)
tp_belt_rf <- predict(mdl_belt_rf, newdata=training)
tp_arm_rf <- predict(mdl_arm_rf, newdata=training)
tp_dumbbell_rf <- predict(mdl_dumbbell_rf, newdata=training)
tp_forearm_rf <- predict(mdl_forearm_rf, newdata=training)

train_ensemble <- cbind.data.frame(classe=tconfirm,belt_rf=tp_belt_rf,arm_rf=tp_arm_rf,dumbbell_rf=tp_dumbbell_rf,forearm_rf=tp_forearm_rf)

set.seed(897)

mdl_ensemble <- train(classe~.,
                   train_ensemble,
                   trControl=train_control,
                   method = "nb")

print(mdl_ensemble)
```

```{r confmat, echo=FALSE, message=FALSE, warning=FALSE}

confusionMatrix(mdl_ensemble, norm = "none", dnn=c("Prediction","Reference"))

```

```{r confmat, echo=FALSE, message=FALSE, warning=FALSE}

test_belt_rf <- predict(mdl_belt_rf, newdata=testing)
test_arm_rf <- predict(mdl_arm_rf, newdata=testing)
test_dumbbell_rf <- predict(mdl_dumbbell_rf, newdata=testing)
test_forearm_rf <- predict(mdl_forearm_rf, newdata=testing)
test_ensemble <- cbind.data.frame(belt_rf=test_belt_rf,arm_rf=test_arm_rf,dumbbell_rf=test_dumbbell_rf,forearm_rf=test_forearm_rf)

trainingpred <- predict(mdl_ensemble, test_ensemble)
trainingpred

```